---
title: "Day2Data"
author: "Robin"
date: "3/13/2018"
output: html_document
---
#Setup
```{r}
library(tidyverse)
library(skimr)
library(ggplot2)
```
Order loading libraries matters because they will mask 'override' previous. Can specify which package to use later
```{r}
gapminder <- read_csv("https://raw.githubusercontent.com/jules32/2017-11-30-MBARI/gh-pages/data/gapminder.csv")
```

```{r}
head(gapminder, 2)
tail(gapminder, 10)
```

```{r}
str(gapminder)
```

```{r}
dim(gapminder)
ncol(gapminder)
nrow(gapminder)
```
$ is a way to select column out of dataframe - will give you the choices
```{r}
unique(gapminder$continent)
unique(gapminder[,"continent"])
```

```{r}
summary(gapminder)
skim(gapminder)
```
```{r}
filter()
stats::filter
dplyr::filter(gapminder, country == "Switzerland")
dplyr::filter(gapminder, lifeExp < 30)
filter(gapminder, country == "Afghanistan", year == max(year))
```
Specify package you want to use i.e. dplyr::
```{r}
dplyr::select(gapminder, country, year, pop)
dplyr::select(gapminder, -continent)
gapminder_no_cont <- dplyr::select(gapminder, -continent)
dplyr::filter(gapminder_no_cont, year == 2007)
```
Pipe %>% - take result from first and use it as data for next to get next result without making new variables.
```{r}
dplyr::select(gapminder, -continent) %>% 
  filter(year == 2007)
```
Mutate
```{r}
gapminder %>%
  mutate(gdp = gdpPercap * pop) %>%
  arrange(desc(gdp)) %>%
  filter(year == 2007)
```

Grouping
```{r}
gapminder %>%
  group_by(country) %>%
  mutate(gdp = gdpPercap * pop, max(gdp)) %>%
  ungroup()
```

Group and Summarize
```{r}
gapminder %>%
  group_by(country) %>%
  mutate(gdp = pop * gdpPercap) %>%
  summarise(max_gdp = max(gdp)) %>%
  arrange(desc(max_gdp))
```

## Joining Datasets, left (first from pipe) + right (second from pipe)
left_join - keep key from left table and join to right
right_join
inner_join - retain only rows in both sets
outer_join - retain all values, all rows (must have by = "x1" to make sure combine)
```{r}
co2 <- readr::read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/co2.csv")
head(co2)
summary(co2)
skim(co2)
```
```{r}
gap_2007 <- gapminder %>%
  filter(year == 2007) 
```
Do checking with anti_join or setdiff
```{r}
left_join(gap_2007, co2, by = "country")
right_join(gap_2007, co2, by = "country")
```

## tidyr
```{r}
AirPassengers
```
```{r}
gap_wide <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder_wide.csv")
```
Use Fn + F1 for help on function while you cursor is in the text of the function
use convert = TRUE while separating to try to get correct data format for output table.
```{r}
gap_better <- gap_wide %>%
  tidyr::gather(observation, observ_value, -continent, -country) %>%
  separate(observation, into = c("obs_type", "year"), sep = "_", convert = TRUE)

```

#Partnering

